NLTK: The Natural Language Toolkit

### 2002-03-28 technical/acl2002/submission.tex
- Converted "Evaluation" section to text - Converted "Conclusion" section to prose - Minor changes
svn/trunk@523

% Outline:
%   - abstract
%   - introduction
%   - choice of programming language
%   - design criteria
%   - modules
%   - uses
%     - assignments/projects
%     - class demonstrations
%   - evaluation
%   - other approaches
%   - conclusion

@@

% Steven: do you want to add anything here?  Say anything about HMMS? 
% Will you have implemented more interfaces by the time of ACL?
The \texttt{fsa} module defines a data type for encoding finite state
automata; and an interface for creating automata from
regular expressions.

@@

A set of visualization modules define graphical interfaces for viewing
and editing data structures; and graphical tools for experimenting
with tasks.  The \texttt{draw.tree} module provides a simple graphical
interface for displaying tree structures; and the
\texttt{draw.tree\_edit} module provides an interface for building and
modifying tree structures.  The \texttt{draw.plot\_graph} module can be
used to graph mathematical functions.  The \texttt{draw.fsa} module
provides a graphical tool for displaying and simulating finite state
automata.  The \texttt{draw.chart} module provides an interactive
graphical tool for experimenting with chart parsers.

@@

% [EL] This makes it sound like we GRADED them using their precision,
% recall, and f-measure.  Either say something about the contest, or
% leave this out:
Each student's project was scored using its precision, recall, and
F-measure.

@@

This tool can be used to explain several different aspects of chart
parsing.  First, it can be used to explain the basic chart data
structure, and to show how edges can represent hypotheses about
syntactic constituents.  It can then be used to demonstrate and
explain the individual rules that the chart parsers uses to create new
edges.  Finally, it can be used to show how these individual rules
combine to find a complete parse for a given sentence.

@@

About CIS-530: mixture of CS and linguistics students.
Approx 40 hours of classes.  All class materials (slides,
assignments) are posted on the course website
\url{http://www.cis.upenn.edu/~cis530/}.

The Language: Python was new to 95\% of students.  Some were initially
bothered about its block structuring, which is sensitive to white
space.  All quickly learnt the language, and reported that any initial
reservations were unfounded.  Many students said that they enjoyed
writing Python code.

CD-ROM: put NLTK and data on CD-ROM and gave it out at the start of class.

The toolkit: students liked being able to do interesting projects from
the outset, liked all the documentation, liked being able to run
everything on their machine at home, ...

Problems: available corpus data, ...

Other strengths / weaknesses

@@

\textit{This section will include a discussion of other approaches to
implementing computational linguistics courseware, including:
Programming for Linguistics: Java Technology for Language
Researchers\cite{Hammond02}; The OpenNLP project\cite{opennlp};
others?}

% ===================== Conclusion =====================
\section{Conclusion}
\label{sec:conclusion}

Overview of the paper.

NOTES: Many CL textbooks provide good exercises at the end of each
chapter.  This isn't enough for many situations.  E.g. students spend
lots of time writing interface code (e.g. to process a corpus).
E.g. teachers spend lots of time creating do-able assignments...

NLTK is an open source project, and we welcome any contributions.  We
deliberately structured NLTK to facilitate parallel development.
Readers who are interested in contributing to NLTK, or who have any
suggestions for improvements, are encouraged to contact the authors.

% include URL (a second time; 1st was in the intro) in conclusion

>>>>

% Outline:
%   - abstract (summary, above.)
%   - introduction
%   - choice of programming language
%   - design criteria
%   - modules
%   - uses
%     - assignments/projects
%     - class demonstrations
%   - evaluation
%   - other approaches
%   - conclusion

@@

% [EL] Steven, do you want to add anything here?  Say anything about HMMS? 
% Will you have implemented more interfaces by the time of ACL?
The \texttt{fsa} module defines a data type for encoding finite state
automata; and an interface for creating automata from regular
expressions.

@@

% jb: this is broken up oddly.
A set of visualization modules define graphical tools for viewing and
editing data structures, and for experimenting with NLP tasks.  The
\texttt{draw.tree} module provides a simple graphical interface for
displaying tree structures; and the
\texttt{draw.tree\_edit} module provides an interface for building and
modifying tree structures.  The \texttt{draw.plot\_graph} module can be
used to graph mathematical functions.  The \texttt{draw.fsa} module
provides a graphical tool for displaying and simulating finite state
automata.  The \texttt{draw.chart} module provides an interactive
graphical tool for experimenting with chart parsers.

@@

%% [EL] This makes it sound like we GRADED them using their precision,
%% recall, and f-measure.  Either say something about the contest, or
%% leave this out:
%Each student's project was scored using its precision, recall, and
%F-measure.

@@

This tool can be used to explain several different aspects of chart
parsing.  First, it can be used to explain the basic chart data
structure, and to show how edges can represent hypotheses about
syntactic constituents.  It can then be used to demonstrate and
explain the individual rules that the chart parser uses to create new
edges.  Finally, it can be used to show how these individual rules
combine to find a complete parse for a given sentence.

@@

% Do we want to mention that there were 40h of lecture time?
% Should we be more explicit about who took the class?
We used NLTK as a basis for the assignments and student projects in
CIS-530, an introductory computational linguistics class taught at the
University of Pennsylvania.  CIS-530 is a graduate level class,
although some advanced undergraduates were also enrolled.  Most
students had either a computer science background or a linguistics
backgroud.  Students were required to complete five assignments, two
exams, and a final project.  All class materials are available from
the corse website \url{http://www.cis.upenn.edu/~cis530/}.

% (Leaving this out for now\ldots)
%The Language: Python was new to 95\% of students.  Some were initially
%bothered about its block structuring, which is sensitive to white
%space.  All quickly learnt the language, and reported that any initial
%reservations were unfounded.  Many students said that they enjoyed
%writing Python code.

% Where does this go??
%At the beginning of the semester, we distributed 2 CD-ROMs to the
%students, containing NLTK and all of the corpera that we would use for
%the class.

The experience of using NLTK was very positive, both for us and for
the students.  The students liked the fact that they could do
interesting projects from the outset.  They also liked being able to
run everything on their computer at home.  The students found the
extensive documentation very helpful for learning to use the toolkit.
They found the interfaces defined by NLTK intuitive, and appreciated
the ease with which they could combine different components to create
complete NLP systems.

We did encounter a few difficulties during the semester.  One problem
is finding large, clean, open-source corpera that the students can use
for their assignments.  This became especially apparent when students
started planning their final projects, since many projects required
different types of corpera.  Another issue was the fact that we were
actively developing NLTK during the semester; some modules were only
completed one or two weeks before we used them.  As a result, students
who worked at home needed to download new versions of the toolkit
several times throughout the semester.  Luckily, Python has extensive
support for installation scripts, which made these upgrades much
simpler.  The students encountered a couple of bugs in the toolkit,
but none were serious, and all were quickly corrected.

%Problems: available corpus data, ...
%
%Other strengths / weaknesses

@@

\textit{This section will include a discussion of other approaches to
implementing computational linguistics courseware, including
\emph{Programming for Linguistics: Java Technology for Language
Researchers}\cite{Hammond02} and \emph{the OpenNLP project}\cite{opennlp}.}

% ===================== Conclusion =====================
\section{Conclusions and Future Work}
\label{sec:conclusion}

%Overview of the paper:
% [EL] Pick better 3 adjectives :)
NLTK provides a simple, extensible, uniform framework for student
assignments and class demonstrations.  It is well documented, easy to
learn, and simple to use.  We hope that NLTK will allow computational
linguistics classes to include more hands-on experience building NLP
components and systems.

% [EL] No room for this for now\ldots
%NOTES: Many CL textbooks provide good exercises at the end of each
%chapter.  This isn't enough for many situations.  E.g. students spend
%lots of time writing interface code (e.g. to process a corpus).
%E.g. teachers spend lots of time creating do-able assignments...

%Future work: 
We plan to continue extending the breadth of materials covered by the
toolkit.  We are currently working on NLTK modules for Hidden Markov
Models, probablistic parsing, and tree adjoining grammars.  We also
plan to increase the number of algorithms implemented by some existing
modules, such as the text classification module.  Finally, we are
putting together a standard collection of large, clean, open-source
corpera, including corpera appropriate for every module defined by the
toolkit.

NLTK is an open source project, and we welcome any contributions.
Readers who are interested in contributing to NLTK, or who have
suggestions for improvements, are encouraged to contact the authors.

% include URL (a second time; 1st was in the intro) in conclusion??
